\chapter[JFeature: Know Your Corpus]{\texorpdfstring{%
JFeature: Know Your Corpus}{%
JFeature: Know Your Corpus}}

\paperRemark{\paperIIref}
\paper{paper:jfeature}

\section{Abstract}
Software corpora are crucial for evaluating research artifacts and ensuring repeatability of outcomes. Corpora such as DaCapo and Defects4J provide
a collection of real-world open-source  projects for evaluating the robustness and performance of software tools like static analysers.
\emph{However, what do we know about these corpora? What do we know about their composition? Are they really suited for our particular problem?}
We developed \textsc{JFeature}, an extensible static analysis tool that extracts syntactic and semantic features from Java programs, to assist developers in answering these questions.
We demonstrate the potential of  \textsc{JFeature} by applying it to four widely-used corpora in the program analysis area, and we suggest other applications, including
 longitudinal studies of individual Java projects and the creation of new corpora.


% Source-Code Analysis, Software Tools, Software Corpora


\section{Introduction}%
\label{JF:sec:introduction}
\input{papers/jfeature/sec1-introduction.tex}

\section{JFeature: automatic feature extraction}%
\label{JF:sec:jfeature}
\input{papers/jfeature/sec2-jfeature.tex}

\section{Corpora Analysis}%
\label{JF:sec:corpora}
\input{papers/jfeature/sec3-corpora-analysis.tex}

\section{Extensibility}%
\label{JF:sec:extension}
\input{papers/jfeature/sec4-extension.tex}

\section{Use cases for JFeature}%
\label{JF:sec:discussion}
\input{papers/jfeature/sec5-discussion.tex}

\section{Related work}%
\label{JF:sec:related-works}
\input{papers/jfeature/sec6-related-works.tex}

\section{Conclusions}%
\label{JF:sec:conclusions}
\input{papers/jfeature/sec7-conclusions.tex}


\section*{Acknowledgements}

This work was partially supported by the Wallenberg AI, Autonomous Systems and Software Program (WASP) funded by the Knut and Alice Wallenberg Foundation.

{\raggedright%
\printbibliography[segment=\therefsegment,heading=subbibliography]
}